---
title: "Week-1-Intro"
author: "Madeline Carlson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

```{r data}
data("penguins")
penguins
```

```{r explore data}
class(penguins)
str(penguins)

class(penguins$species)
str(penguins$species)
```

```{r lists}
myList <- list("apple", 1993, FALSE, penguins)
str(myList)

list(myList, list("more stuff here", list("and more")))

names(myList) <- c("fruit", "year", "logic", "data")
names(myList)
```

```{r indexing}
myList[[1]]
myList[["data"]]

myVector <- c("apple", "banana", "pear")
myVector[2]

penguins[1:5, 2]
penguins[1:5, "island"]
penguins[1, 1:5]
penguins[1:5, c("species","sex")]
penguins[penguins$sex=='female',]

penguins$species
```

# Exercise

1.  Why don't the following lines of code work? Tweak each one so the code runs

```{r 1}
myList[["fruit"]]

penguins$flipper_length_mm

penguins[penguins$island=='Dream',]
```

2.    How many species are in the penguins dataset? What islands were the data collected for? (Note: the unique() function might help)

3.    Use indexing to create a new data frame that has only 3 columns: species, island and flipper length columns, and subset all rows for just the ‘Dream’ island.

4.    Use indexing and the mean() function to find the average flipper length for the Adelie species on Dream island. (Note: explore the mean() function and how to deal with NA values).




