---
title: "Week-1-Spatial"
author: "Madeline Carlson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```


```{r, include=F}
counties <- counties(state = "CO")
roads <- roads(state = "CO", county = "Larimer")
tmap_mode("view")

poudre_hwy <- roads %>% 
  filter(FULLNAME == "Poudre Canyon Hwy")
```

```{r}
poudre_points <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))
poudre_points_sf <- st_as_sf(poudre_points, coords = c("long", "lat"), crs = 4326)

poudre_points_prj <- st_transform(poudre_points_sf, st_crs(counties))
st_crs(poudre_points_prj) == st_crs(counties)
```

```{r, include=FALSE}
elevation <- get_elev_raster(counties, z = 7)

tm_shape(elevation)+
  tm_raster(style = "cont", title = "Elevation (m)")
```

```{r}
elevation <- rast(elevation)
names(elevation) <- "Elevation"

elevation_crop <- crop(elevation, ext(roads))

tm_shape(elevation, bbox = st_bbox(poudre_hwy))+
  tm_raster(style = "cont", title = "Elevation (m)")+
tm_shape(poudre_hwy)+
  tm_lines()+
tm_shape(poudre_points_prj)+
  tm_dots(size = 0.2)
```

# 4. Exercises

## 1. Filter out the counties data set to only include Larimer, Denver, and Pueblo counties.

```{r}
counties3<-counties%>%
  filter(NAME %in% c("Larimer", "Denver", "Pueblo"))
```

## 2. Make a map of the counties data colored by county area. Make a second map of counties colored by their total area of water.

```{r}
tm_shape(counties)+
  tm_polygons(col = "ALAND")

tm_shape(counties)+
  tm_polygons(col = "AWATER")
```

## 3. Make a barplot comparing the elevation of your 3 points in the Poudre Canyon (note: explore the extract() function in the terra package).

```{r}
poudre_points_prj$elevation <- extract(elevation, vect(poudre_points_prj))$Elevation 
    
poudre_points_prj %>% 
  ggplot() +
  geom_bar(aes(x = name, y = elevation), stat = "identity")
```

## 4.  Why are there 4 features in our Poudre Canyon Highway variable instead of 1?

There are 3 small segments of highway not attached to the main highway but have the same name.
