---
title: "Week-1-Intro"
author: "Madeline Carlson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

```{r data, include=F}
data("penguins")
penguins
myList <- list("apple", 1993, FALSE, penguins)
list(myList, list("more stuff here", list("and more")))
names(myList) <- c("fruit", "year", "logic", "data")
```

# Exercise 6.3.1

##1. Why don't the following lines of code work? Tweak each one so the code runs

```{r 1}
myList[["fruit"]]

penguins$flipper_length_mm

penguins[penguins$island=='Dream',]
```

##2. How many species are in the penguins dataset? What islands were the data collected for? (Note: the unique() function might help)

```{r}
unique(penguins$species)
unique(penguins$island)
```

There are 3 different species from the islands of Torgersen, Biscoe and Dream.

##3. Use indexing to create a new data frame that has only 3 columns: species, island and flipper length columns, and subset all rows for just the 'Dream' island.

```{r}
indexing1<-penguins[penguins$island=='Dream', c("species", "island", "flipper_length_mm")]
```

##4. Use indexing and the mean() function to find the average flipper length for the Adelie species on Dream island. (Note: explore the mean() function and how to deal with NA values).

```{r}
indexing2<-indexing1[indexing1$species=='Adelie',]
mean(indexing2$flipper_length_mm)
```

# Exercise 6.3.2

##1. Reorder the variables in penguins so that year is the first column followed by the rest (Hint: look into the use of everything()).

```{r}
penguins<-penguins%>%
  relocate(year, species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex)
#or
penguins%>%
  select(year, everything())
```

##2. Create a new column called 'size_group' where individuals with body mass greater than the overall average are called 'large' and those smaller are called 'small'.

```{r}
body_mass_avg <- mean(penguins$body_mass_g, na.rm = TRUE)
penguins<-penguins %>% 
  mutate(size_group = if_else(body_mass_g>=body_mass_avg,
                              paste('large'), paste('small')))
```

##3. Find out which year for each species were individuals on average the largest according to body mass.

```{r}
penguins %>% 
  group_by(year) %>% 
  summarise(body_mass_avg = mean(body_mass_g, na.rm = TRUE))
```

##4. You want to filter data for years that are not in a vector of given years, but this code doesn't work. Tweak it so that it does. (Yes, you could just filter year to equal 2007 in this case but there is a trouble-shooting lessons here).

```{r}
penguins7<-penguins %>% 
  filter(!year %in% c("2008", "2009"))
```

# Exercise 7.1

##1. Make a barplot showing the average flipper length for each species.

```{r}
penguins%>%
  group_by(species)%>%
  summarise(mean_flipper = mean(flipper_length_mm, na.rm = T))%>%
  ggplot(aes(x = species, y = mean_flipper))+
    geom_bar(stat = "identity", show.legend = F, fill = c("darkorange","darkorchid","cyan4"))+
    theme_bw()+
    labs(x = "Species",
         y = "Flipper Length (mm)")
```

##2. Make a scatter plot of bill length compared to bill depth but only for observations on the Dream island.

```{r}
ggplot(penguins) +
  geom_point(filter(penguins, island == "Dream"), 
             mapping = aes(x = bill_length_mm, y = bill_depth_mm), color = "cyan3")+
  theme_bw()+
  labs(x = "Bill Length (mm)",
       y = "Bill Depth (mm)")
```
